---
title: "KIND Introductory Shiny"
subtitle: "session 2"
author: "Brendan Clarke"
format: 
    revealjs:
        logo: "..//src//images//KLN_banner_v05_125.png"
        css: "..//src//images//logo.css"
scrollable: true
smaller: true
execute: 
    echo: true
    eval: false
---
## Session goal: make the dashboard more useful

## Session learning outcomes

+ consolidate session one
  + Shiny structure
  + function pairs (`render*()` and `output*()`)
  + input options
+ extend those skills cross `ui()` and `server()` 
  + additional input selections
  + more powerful reactivity
  + more ambitious R code in `server()`

## adding more interactivity

Let's start again with an empty Shiny app. During this session, we'll build this into a mini-dashboard that:

+ uses `ae_attendances`
+ allows us to plot time trends of attendances between organisations
+ includes a national average benchmark
+ and uses several kinds of interactivity

## Milestone 01
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_01.R
```

## Input options

Time for the [Shiny Cheatsheet](https://shiny.rstudio.com/images/shiny-cheatsheet.pdf)

+ `checkboxGroupInput()`
+ `radioButtons()`

Arguments:

+ name
+ label
+ choices
+ selected

## Adding to the UI

Let's add some appropriate elements to our `ui()` for this graph:

+ `checkboxGroupInput()` to allow us to select an org
  + let's start with a simple vector of orgs `c("RJ1", "RLN", "RXK")`
+ `radioButtons()` to select the type

## Milestone 02
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_02.R
```

## Plotting attendance by org

Here's a code snippet to allow us to produce a `ggplot()` based on different orgs and attendance types:

```{!r}
ae_attendances %>%
  filter(org_code %in% c("RJ1", "RLN", "RXK") & type == "1") %>%
  ggplot() +
  geom_line(aes(x = period, y = attendances, color = org_code))
```
+ `%in%` - very useful for filtering on several values

+ Let's try that code now to check the output

## Milestone 03
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_03.R
```


What do we need to add to include a static version of the graph in our dashboard?
+ in `ui()`?
+ in `server()`?

## Milestone 04
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_04.R
```

Note this is not interactive - we've just put the static R code into the `server()`

## Adding interactivity

We need to connect our inputs with the `server()`:

+ `input$org` rather than `c("RJ1", "RLN", "RXK")`
  + setting a default value of "RJ1" using `selected = "RJ1"`
+ `type == input$type_select`

## Milestone 05
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_05.R
```

## A trap: hard-coded values

Typing values for users to select from is tedious. Can we cheat to get R to do the work for us?

+ We can include simple snippets of R code directly inside UI elements

+ (R question): how to get the possible values of `type` directly from `ae_attendances`?
+ (Shiny question): how to include those in our `radioButtons()`?

## Milestone 06
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_06.R
```

We could do something similar for the orgs too, which we'll do later.

## Adding a national benchmark

+ get the R code working
+ add that code to `server()`
+ add `ui()` options
+ test

## Milestone 07
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_07.R
```

## Adding the benchmark code to `server()`

+ We only want the benchmark to update when we switch type
+ to do that, we'll use `reactive()`

To make a `reactive()`:
```{!r}
benchmark <- reactive({
<<code>>
})
```

To use a `reactive()`:

+ special way of calling reactive objects, similar to a function
+ `data = benchmark()` etc

## Milestone 08
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_08.R
```

Lots going on here!

## `reactive()`

+ `reactive()` makes a reactive `benchmark()` object summarising the national data
  + inside `server()`
  + gets updated only when `type` changes
+ then use that `benchmark()` in the `ggplot`
  + special way of calling reactive objects, similar to a function
  + `data = benchmark()` etc

## Additional tweaks

+ set-up a vector of key orgs where data is available for all the months and add to UI using checkboxGroupInput

```{r}
key_orgs <- ae_attendances %>%
  group_by(period, org_code) %>%
  add_tally(name = "type_tally") %>%
  filter(type_tally == 3) %>%
  ungroup() %>%
  group_by(org_code, type) %>%
  add_count(name = "months") %>%
  filter(months == 36) %>%
  pull(org_code) %>%
  unique()
```
+ `ggplotly()`, `renderPlotly()`, and `plotlyOutput()`

## Milestone 09
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_09.R
```

## Finishing touches

Conditional execution for our benchmark

  + add `checkboxInput()` to `ui()`
  + add a new `reactive()` object from that input
  + use that new `reactive()` in an `if`/`else`
    + in the `TRUE` arm of that `if`/`else`, add the benchmarked `ggplot`
    + in the `FALSE` arm of that `if`/`else`, add the non-benchmarked `ggplot`

## Milestone 10
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s02_ms_10.R
```

## Next time...
+ outsourcing more complicated R code to a script
+ thinking about UI design





  
 <!-- all milestones included -->
