---
title: "KIND Introductory Shiny"
subtitle: "session 4"
author: "Brendan Clarke"
format: 
    revealjs:
        logo: "..//src//images//KLN_banner_v05_125.png"
        css: "..//src//images//logo.css"
scrollable: true
smaller: true
execute: 
    echo: true
    eval: false
---
## Session learning outcomes

+ `shinydashboard` menu and sections
+ adding those new functions as pages in our dashboard
+ tweaks, tidying up, and story-telling
+ thinking about project architecture

## `shinydashboard` menu and sections
  
+ Let's look again at the `ui()` that we had at the end of the last session

## Milestone 01
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s04_ms_01.R
```

## `shinydashboard` `ui()`
+ we'll add some elements to the page to allow us to show off the other graphs we drew last time. First, add some content to `dashboardHeader()`
  + add a `title = "whatever you like"`
  + optional `titleWidth` argument to preserve long titles

## `dashboardSidebar()`
+ set a `width =`, ideally corresponding with the `titleWidth` value from `dashboardHeader`
+ `collapsed` TRUE/FALSE *au choix*
+ `sidebarMenu()`
  + 5 x `menuItem()`s, each containing:
    + a tab title, for display
    + `tabName = ` tab name
    + `icon = icon("icon name")`
    
## Icons...
+ really useful for dashboards, but poorly documented
+ simple cases (like this one), supply a fontawesome icon name inside `icon()`
  + e.g. `icon("info-circle")` becomes 

```{r}
#| echo: false
#| eval: true
shiny::icon("info-circle")
```



## Milestone 02
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s04_ms_02.R
```

## Milestone 03
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s04_ms_03.R
```

## Milestone 04
```{r}
#| echo: true
#| eval: false
#| file: ../milestones/shiny intro_s04_ms_04.R
```

  
 <!-- all milestones included -->
